
\section{Conclusion}
\label{sec:conclusion}
\method is a unified model designed to understand and generate multiple modalities, including text and real-time speech. To enhance video integration, we've introduced a new positional embedding method called TMRoPE, which aligns audio and video timing. Our Thinker-Talker framework supports real-time speech generation while minimizing interference across different modalities. Additionally, we employ techniques such as block-wise audio/vision encoding and a sliding window mechanism for code-to-wav generation. This innovative model excels in complex audio-visual interactions and emotional context in speech dialogues. Comprehensive evaluations show that \method outperforms similarly sized single-modality models, particularly in following voice commands, and achieves state-of-the-art performance in multi-modal tasks.

In the development of the model, we have identified several critical issues that have often been overlooked by researchers in previous academic studies, such as video OCR and audio-video collaborative understanding. Addressing these challenges necessitates collaboration between the academic and industrial sectors, particularly in building comprehensive evaluation benchmarks and research datasets. We believe \method represents a significant advancement toward artificial general intelligence (AGI). Our future goals include developing a more robust and faster model with expanded output capabilities across various modalities like images, videos, and music. 

