\section{Pre-training}
\method consists of three training stages. In the first stage, we lock the LLM parameters and focus exclusively on training the vision encoder and audio encoder, utilizing a vast corpus of audio-text and image-text pairs to enhance semantic understanding within the LLM. In the second stage, we unfreeze all parameters and train with a wider range of multimodal data for more comprehensive learning. In the final stage, we use data with a sequence length of 32k to enhance the model's ability to understand complex long-sequence data.

The model is pre-trained on a diverse dataset that includes various types such as image-text, video-text, video-audio, audio-text and text corpus. We replace the hierarchical tags with the natural language prompts following Qwen2-Audio~\citep{qwen2audio}, which can improve better generalization ability and better instruction following ability.

During the initial pre-training phase, the LLM component of \method is initialized using the parameters from Qwen2.5~\citep{qwen2.5}, while the vision encoder is the same as Qwen2.5-VL, and the audio encoder is initialized with the Whisper-large-v3~\citep{Whisper}. The two encoders are trained separately on the fixed LLM, with both initially focusing on training their respective adapters before training the encoders. This foundational training is crucial in equipping the model with a robust understanding of core visual-textual and audio-textual correlations and alignments.

The second phase of pre-training marks a significant advancement by incorporating an additional 800 billion tokens of image and video related data, 300 billion tokens of audio related data, and 100 billion tokens of video with audio related data. This phase introduces a larger volume of mixed multimodal data and a wider variety of tasks, which enhances the interaction and deepens the understanding between auditory, visual, and textual information. The inclusion of multimodal, multitask datasets is crucial for developing the model's ability to handle multiple tasks and modalities simultaneously, a vital capability for managing complex real-world datasets. Moreover, pure text data plays an essential role in maintaining and improving language proficiency.

To improve training efficiency, we limited the maximum token length to 8192 tokens in the previous stages. Then, we incorporat long audio and long video data and extend the original text, audio, image, and video  data to 32,768 tokens for training. Experimental results indicate that our data shows significant improvement in supporting long sequence data


