\documentclass{article}

\usepackage{microtype}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{xspace}

\usepackage{hyperref}
\newcommand{\theHalgorithm}{\arabic{algorithm}}
\usepackage[preprint]{icml2026}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage{amsthm}
\usepackage[capitalize,noabbrev]{cleveref}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% THEOREMS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{assumption}[theorem]{Assumption}
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}

\usepackage[textsize=tiny]{todonotes}

% Miscellaneous new commands
\newcommand{\ours}{Hibiki-Zero\xspace}
\newcommand{\red}{\textcolor{red}}
\newcommand{\ensuremathmode}[1]{%
  \ifmmode
    #1% Already in math mode, use as is
  \else
    $#1$% Not in math mode, wrap with dollar signs
  \fi
}
\newcommand{\softmax}{\mathrm{softmax}}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\E}{\mathbb{E}}
\newcommand{\proba}[1]{\mathbb{P}\left[#1\right]}
\newcommand{\waveform}{\ensuremathmode{x}}
\newcommand{\dims}{D}
\newcommand{\numsamples}{L}
\newcommand{\numtokens}{S}
\newcommand{\numquantizers}{\ensuremathmode{Q}}
\newcommand{\cardaudio}{\ensuremathmode{N_A}}
\newcommand{\cardtext}{\ensuremathmode{N_W}}
\newcommand{\acousticdelay}{\ensuremathmode{\tau}}
\newcommand{\yescheck}{\ding{51}}
\newcommand{\nocheck}{\ding{55}}

\begin{document}

\twocolumn[
\icmltitle{Simultaneous Speech-to-Speech Translation Without Aligned Data}

\icmlsetsymbol{equal}{*}

\begin{icmlauthorlist}
\icmlauthor{Tom Labiausse}{kyutai}
\icmlauthor{Romain Fabre}{kyutai}
\icmlauthor{Yannick Est\`eve}{lia}
\icmlauthor{Alexandre D\'efossez}{kyutai,gradium}
\icmlauthor{Neil Zeghidour}{gradium}
\end{icmlauthorlist}

\icmlaffiliation{kyutai}{Kyutai, Paris, France}
\icmlaffiliation{gradium}{Gradium, Paris, France}
\icmlaffiliation{lia}{LIA, University of Avignon, France}

\icmlcorrespondingauthor{\\ \ours{}}{hibiki@kyutai.org}

\vskip 0.3in
]

\printAffiliationsAndNotice{}  % no special notice (required even if empty)

\begin{abstract}
Simultaneous speech translation requires translating source speech into a target language in real-time while handling non-monotonic word dependencies. Traditional approaches rely on supervised training with word-level aligned data, which is difficult to collect at scale and thus depends on synthetic alignments using language-specific heuristics that are suboptimal. We propose \ours{}, which eliminates the need for word-level alignments entirely. This fundamentally simplifies the training pipeline and enables seamless scaling to diverse languages with varying grammatical structures, removing the bottleneck of designing language-specific alignment heuristics. We first train on sentence-level aligned data to learn speech translation at high latency, then apply a novel reinforcement learning strategy using GRPO to optimize latency while preserving translation quality. \ours{} achieves state-of-the-art performance in translation accuracy, latency, voice transfer, and naturalness across five X-to-English tasks.  Moreover, we demonstrate that our model can be adapted to support a new input language with less than 1000h of speech. We provide examples, model weights, inference code\footnote{\href{https://github.com/kyutai-labs/hibiki-zero}{github.com/kyutai-labs/hibiki-zero}} and we release a benchmark containing 45h of multilingual data for speech translation evaluation.\footnote{\href{https://huggingface.co/collections/kyutai/hibiki-zero}{huggingface.co/collections/kyutai/hibiki-zero}}
\end{abstract}

\input{sections/introduction}
\input{sections/related}
\input{sections/method}
\input{sections/experiments}
\input{sections/conclusion}

\bibliography{references}
\bibliographystyle{icml2026}

\newpage
\appendix
\onecolumn
\input{sections/appendix}

\end{document}