\section{Introduction}
\label{introduction}

\looseness=-1
We introduce \ours{}, a system for simultaneous and expressive speech-to-speech (S2ST) and speech-to-text (S2TT) translation that does not require aligned data for training. Unlike offline speech translation systems that access the full source utterance before translating, simultaneous translation must produce output incrementally while maintaining both translation accuracy and speech naturalness. This requires learning a fine-grained translation policy that determines when to listen and when to speak. The most straightforward approach to learning such a policy is through supervised training on aligned data. However, human interpretation data with word-level alignments is virtually non-existent, forcing state-of-the-art systems to rely on synthetic data with automatic alignments~\cite{hibiki}. These automatic alignments are inherently limited, as they depend on hand-crafted heuristics rather than being learned from data.

\looseness=-1

\ours{} is a decoder-only model that synchronously receives source speech and generates translated speech leveraging a multistream architecture originally introduced by~\citet{moshi}. Unlike Hibiki~\cite{hibiki}, \ours{} is not trained with supervised learning on synthetic interpretation data but rather casts joint optimization of translation quality and latency as a reinforcement learning (RL) problem. While we still require a base model before the RL phase, it is trained using sentence-level aligned data which can be more easily constructed independently of the language compared to word-level aligned data. During RL, we exploit the sentence-level aspect of our data to design a simple reward system based on BLEU score~\cite{bleu} only. To achieve this, we compute rewards at multiple intermediate instants during the translation of an input speech utterance by leveraging the simultaneous text translation also produced by our model. Using these \textit{process rewards}, we obtain fine-grained local \textit{advantages} across multiple translations from the same input. We then adapt GRPO~\cite{grpo} to our multistream architecture, using these advantages to optimize the model.

\looseness=-1
In a multilingual-to-English translation task, \ours{} outperforms previous state-of-the-art work in translation quality, latency, speaker identity preservation, and speech naturalness. We also retain all the benefits of multistream modeling such as batching and real-time inference on GPU while removing the necessity to build interpretation-like training data thus considerably simplifying the development of such models. We even demonstrate that \ours{} can adapt to a new input language with less than 1000h of training data marking an important step to make high quality speech translation (ST) available in more languages. We will release our data preparation code, model weights as well as a 45h multilingual speech benchmark for ST evaluation. 